---
title: "Synth Control"
output: html_document
date: "2023-04-11"
---

```{r}

library(dplyr)
library(ggplot2)
library(lubridate)
library(SCtools)
library(Synth)
theme_set(theme_minimal())
```

```{r}
Full_Data <- readRDS("Full_Data.rds")

```


```{r NO2-graph, fig.cap="Trend of NO2 in Berlin"}

Full_Data %>% 
  filter(AQ_city == "Berlin") %>% 
  group_by(month = month(date_and_time), year = year(date_and_time)) %>% 
  summarize(mean_NO2 = mean(AQ_NO2, na.rm = TRUE)) %>% 
  ggplot(aes(x = as.Date(paste(year, month, "01", sep = "-")), y = mean_NO2)) +
  geom_line() +
  geom_vline(xintercept = as.Date("2022-09-27"), color = "orange") +
  scale_x_date(date_labels = "%b %Y") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y")

```

```{r fig.cap="Trends in Predictors in NO2"}
Full_Data %>% 
  filter(AQ_city == "Berlin") %>% 
  select(FP_diesel_avg, W_wind_speed, W_precipitation, date_and_time) %>% 
  tidyr::gather(predictor, value, -date_and_time) %>% 
  mutate(month = month(date_and_time), year = year(date_and_time)) %>% 
  group_by(predictor, month, year) %>% 
  summarize(mean_value = mean(value, na.rm = TRUE)) %>% 
  ggplot(aes(x = as.Date(paste(year, month, "01", sep = "-")), y = mean_value)) +
  geom_line() +
  facet_wrap(~predictor, scales = "free") +
  coord_cartesian(ylim = c(0, NA))
```

```{r}

# Create a numeric identifier for each city
Full_Data <- Full_Data %>%
  mutate(city_id = as.numeric(factor(AQ_city, levels = unique(AQ_city))))

# Create vector of control city ids
control_cities <- as.integer(c(1, 3:10))

# Create an ungrouped list of comparison city names
comparison_cities <- 
unlist(
  Full_Data %>% 
  group_by(AQ_city) %>% 
  filter(AQ_city != "Berlin") %>% 
  summarize(AQ_city) %>% 
  distinct()
)


```


```{r}

# Transforming months to integer values and group data on a monthly granularity

Full_Data2 <- Full_Data %>% 
  filter(!is.na(date_and_time)) %>%  # Filter out rows with missing date_and_time values
  mutate(city_id = as.numeric(factor(AQ_city, levels = unique(AQ_city))),
         month_year = floor_date(date_and_time, unit = "month"),
         month_factor = as.integer(as.factor(dense_rank(month_year)))) %>% 
  group_by(city_id, AQ_city, month_year, month_factor) %>% 
  summarize(mean_NO2 = mean(AQ_NO2, na.rm = TRUE),
            mean_FP = mean(FP_diesel_avg, na.rm = TRUE),
            mean_WS = mean(W_wind_speed, na.rm = TRUE),
            mean_T = mean(W_air_temperature, na.rm = TRUE)
            )


```

```{r}

# Using my monthly grouped with integer levels

dataprep.out<-dataprep(
	    foo = as.data.frame(Full_Data2),
	    predictors = c("mean_FP", "mean_WS", "mean_T"),
	    predictors.op = "mean",
	    dependent = "mean_NO2",
	    unit.variable = "city_id",
	    time.variable = "month_factor",
	    treatment.identifier = 2,
	    controls.identifier = control_cities,
	    time.predictors.prior = c(1:14),
	    time.optimize.ssr =  c(1:14),
	    special.predictors = list(
      list("mean_NO2", 1:14,"mean")),
	    unit.names.variable = "AQ_city",
	    time.plot = 1:20
	)

```

```{r eval=FALSE}
out <- synth(dataprep.out, Sigf.ipop = 3)
```

```{r}

solution <- out$solution.w %>% 
	as.data.frame()

solution$city_id <- rownames(solution)

solution %>% 
	mutate(city_id = as.numeric(city_id)) %>% 
	left_join(Full_Data4 %>% 
	select(AQ_city,city_id) %>% 
	filter(AQ_city %in% comparison_cities) %>% 
		distinct(), by = "city_id") %>% 
	ggplot(aes(reorder(AQ_city, w.weight), w.weight))+
	geom_col()+
	coord_flip()+
	labs(
		title = "Donor Weights by City",
		y = NULL,
		x = "Weight"
	)
```




```{r fig.cap="Synthetic Russia vs Actual Russia for Alcohol Consumption per Capita"}
path.plot(synth.res = out, dataprep.res = dataprep.out, 
					Xlab = "NO2 patterns", )
```


```{r}
delta_out <- (dataprep.out$Y1plot - (dataprep.out$Y0plot %*% out$solution.w)) %>% 
	as.data.frame()

delta_out$year <- rownames(delta_out)

delta_out%>% 
	knitr::kable(caption = "Difference in NO2 Between Synthetic and Actual Berlin", digits = 1)
```



```{r eval=FALSE}
placebo <- generate.placebos(dataprep.out = dataprep.out,
                             synth.out = out, strategy = "multicore")
```

Now that we have our placebo object, we can represent it graphically with `plot_placebos`. Here we can see our donors (control unit) and our actual treated group. 

```{r fig.cap="Placebo Plot for Control Units"}
plot_placebos(placebo)
```


Equally important we can test the Mean Squared Prediction Error (MSPE). Additionally, because we generated the controls for the placebos, we can see how extreme our values are and thus generate a pseudo p-value to see if our results are significant. Now, it should be mentioned that this test is relatively underpowered as it looks only at the donor states.

```{r}
test_out <- mspe.test(placebo)
test_out$p.val
```

Why look at single number summary, when you can look at a plot? With `mspe.plot` we can visualise the ratios of MSPE for each donor. Here we see very little difference in Russia, which supports our higher p-value. 

```{r fig.cap="Mean Squared Prediction Error Ratios for Russia and Donors"}
mspe.plot(tdf = placebo)
```






```{r}

# Using as.Date on monthly groupings

dataprep.out<-dataprep(
	    foo = as.data.frame(Full_Data4),
	    predictors = c("mean_FP"),
	    predictors.op = "mean",
	    dependent = "mean_NO2",
	    unit.variable = "city_id",
	    time.variable = "month_year",
	    treatment.identifier = 2,
	    controls.identifier = control_cities,
	    time.predictors.prior = c(as.Date("2021-09-01"):as.Date("2022-10-01")),
	    time.optimize.ssr =  c(as.Date("2021-09-01"):as.Date("2022-06-01")),
	    special.predictors = list(
      list("AQ_NO2", as.Date("2021-09-01"):as.Date("2022-06-01") ,"mean")),
	    unit.names.variable = "AQ_city",
	    time.plot = as.Date("2021-09-01"):as.Date("2023-03-31")
	)

```



```{r}

# Using precise dates as time stamps

dataprep.out<-dataprep(
	    foo = as.data.frame(Full_Data3),
	    predictors = c("FP_diesel_avg",
	                   "W_wind_speed",
	                   "W_precipitation"),
	    predictors.op = "mean",
	    dependent = "AQ_NO2",
	    unit.variable = "city_id",
	    time.variable = "date_and_time",
	    treatment.identifier = 2,
	    controls.identifier = control_cities,
	    time.predictors.prior = c(as.Date("2021-09-01"):as.Date("2022-09-27")),
	    time.optimize.ssr =  c(as.Date("2021-09-01"):as.Date("2022-06-01")),
	    special.predictors = list(
      list("AQ_NO2", as.Date("2021-09-01"):as.Date("2022-06-01") ,"mean")),
	    unit.names.variable = "AQ_city",
	    time.plot = as.Date("2021-09-01"):as.Date("2023-03-31")
	)

```

```{r}

# Using month-year as time stamps

Full_Data2$month_year <- paste(month(Full_Data2$date_and_time), year(Full_Data2$date_and_time), sep = "-")

dataprep.out<-dataprep(
	    foo = as.data.frame(Full_Data2),
	    predictors = c("FP_diesel_avg",
	                   "W_wind_speed",
	                   "W_precipitation"),
	    predictors.op = "mean",
	    dependent = "AQ_NO2",
	    unit.variable = "city_id",
	    time.variable = "month_year",
	    treatment.identifier = 2,
	    controls.identifier = control_cities,
	    time.predictors.prior = c("9-2021":"9-2022"),
	    time.optimize.ssr =  c("9-2021":"9-2022"),
	    special.predictors = list(
      list("AQ_NO2", "9-2021":"9-2022" ,"mean")),
	    unit.names.variable = "AQ_city",
	    time.plot = "9-2021":"3-2023"
	)

```

```{r}

# Using a modified condensed data frame so that each city has only one value per hour of NO2

Full_Data3 <- Full_Data2 %>% 
  select(AQ_city, date_and_time, AQ_NO2, city_id, FP_diesel_avg, W_wind_speed, W_precipitation) %>% 
  group_by(AQ_city, city_id, date_and_time) %>% 
  summarise(mean_NO2 = mean(AQ_NO2, na.rm = TRUE),
         mean_FP = mean(FP_diesel_avg, na.rm = TRUE),
         mean_Wspeed = mean(W_wind_speed, na.rm = TRUE),
         mean_Wprec = mean(W_precipitation, na.rm = TRUE)
         )


dataprep.out<-dataprep(
	    foo = as.data.frame(Full_Data3),
	    predictors = c("mean_FP",
	                   "mean_Wspeed",
	                   "mean_Wprec"),
	    predictors.op = "mean",
	    dependent = "mean_NO2",
	    unit.variable = "city_id",
	    time.variable = "date_and_time",
	    treatment.identifier = 2,
	    controls.identifier = control_cities,
	    time.predictors.prior = c(as.Date("2021-09-01"):as.Date("2022-09-27")),
	    time.optimize.ssr =  c(as.Date("2021-09-01"):as.Date("2022-06-01")),
	    special.predictors = list(
      list("mean_NO2", as.Date("2021-09-01"):as.Date("2022-06-01") ,"mean")),
	    unit.names.variable = "AQ_city",
	    time.plot = as.Date("2021-09-01"):as.Date("2023-03-31")
	)

```


```{r}


# Original Data with NAs


# Create a numeric identifier for each city
Full_Data <- Full_Data %>%
  mutate(city_id = as.numeric(factor(AQ_city, levels = unique(AQ_city))))


dataprep.out<-dataprep(
	    foo = as.data.frame(Full_Data),
	    predictors = c("FP_diesel_avg",
	                   "W_wind_speed",
	                   "W_precipitation"),
	    predictors.op = "mean",
	    dependent = "AQ_NO2",
	    unit.variable = "city_id",
	    time.variable = "date_and_time",
	    treatment.identifier = 2,
	    controls.identifier = control_cities,
	    time.predictors.prior = c(as.Date("2021-09-01"):as.Date("2022-09-27")),
	    time.optimize.ssr =  c(as.Date("2021-09-01"):as.Date("2022-06-01")),
	    special.predictors = list(
      list("AQ_NO2", as.Date("2021-09-01"):as.Date("2022-06-01") ,"mean")),
	    unit.names.variable = "AQ_city",
	    time.plot = as.Date("2021-09-01"):as.Date("2023-03-31")
	)


```


```{r}

# Check for any missing dates and times in my data

# Load required libraries
library(lubridate)

# Set the time zone to German time zone
Sys.setenv(TZ = "CET")

# Generate all datetimes in German time zone
all_datetimes <- seq(from = as.POSIXct("2021-09-01 00:00:00", tz = "CET"), 
                     to = as.POSIXct("2023-03-31 23:00:00", tz = "CET"), 
                     by = "hour")

# Extract the datetime values from the dataframe - Replace Data Frame name to check
actual_datetimes <- as.POSIXct(Full_Data2$date_and_time)

# Compare the two sequences and check if there are any missing datetimes
missing_datetimes <- setdiff(all_datetimes, actual_datetimes)

# Check if there are any missing datetimes
if (length(missing_datetimes) == 0) {
  print("There are no missing datetimes")
} else {
  print(paste("There are", length(missing_datetimes), "missing datetimes "))
  print(as.POSIXct(missing_datetimes, origin = "1970-01-01"))
  
  # Fill in data with NAs for missing datetimes
  for (i in missing_datetimes) {
    Full_Data <- rbind(Full_Data, c(date_and_time = i, rep(NA, ncol(Full_Data)-1)))
  }
}


# The only missing values are related to timezone change in Germany

```


